---
title: "AreaCoder"
author: "Connor Wiegand"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: yeti
    highlight: haddock
    toc: yes
    toc_depth: 4
    toc_float: yes
    keep_md: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
## This next line sets the default behaviour for all R chunks in the .Rmd document.
## I recommend you take a look here: https://rmarkdown.rstudio.com/authoring_rcodechunks.html
knitr::opts_chunk$set(eval=T, include=T, echo = T, cache = T, error = TRUE, dpi=300)
```

## Background


## 1) Loading Packages

- Leaflet allows for interactive maps, only in Simple Features (sf)
- Choroplethr uses ggplot and data frames converted from shapefiles (sp)\

```{r libs, cache=F}
## Load your packages here, e.g.
pacman::p_load(rvest, readr, tidyverse, magrittr, data.table, gridExtra, utils)
pacman::p_load(lubridate, zoo)
pacman::p_load(sf, rnaturalearth, rgeos, leaflet, mapproj,)
pacman::p_load(choroplethr, choroplethrAdmin1,choroplethrMaps)
```

## 2) Proof of Concept

Let's start by trying Peru, which can also be found in section 3.1.

### 2-1: Get Telephone Data

Let's start by pulling the area code data that we need from wikipedia:

#### 2-1-0: Read website

```{r proofreadweb}
wiki_site <- read_html("https://en.wikipedia.org/wiki/Telephone_numbers_in_Peru")
```

#### 2-1-1: Scrape for Desired data; format said data
For this, one must copy the xpath via inspect element by hand. In this example, using just the table body (no headers) is what works. This still reads and preserves the headers.

One large challenge for the future is ** *how to automate this?* **
```{r proofxpath}
wiki_xpath <- '/html/body/div[3]/div[3]/div[5]/div[1]/table[2]/tbody'
## Here, using just the table body (no headers) is what works
### headers are still preserved


wiki_wp <- html_element(wiki_site, xpath = wiki_xpath) #*
wiki_table <- html_table(wiki_wp)
```


## 2-2: Get Regions of Peru
The following gives an example choroplethr map for Peru:
```{r choro_ex}
#Works:
peru_regions <- get_admin1_regions("peru")
peru_regions$value = 1:nrow(peru_regions)
admin1_choropleth("peru", peru_regions)

# data(df_pop_state)
# state_choropleth(df_pop_state)

#Doesn't:
data(country.map)
ggplot(country.map, aes(long, lat, group=group)) + geom_polygon()

data("state.map")
ggplot(state.map, aes(long, lat, group=group)) + geom_polygon()

```

```{r}
countries <- ne_countries(returnclass = "sf") %>% st_transform(8857) 
#ggplot(admin1.map, aes(long, lat, group=group)) + geom_polygon()

data(admin1.map)
head(admin1.map)
# data(admin1.regions)
# head(admin1.regions)
admin1_map("peru")
?admin1_map
get_admin1_map("peru")
View(get_admin1_map("peru"))

proof_regions <- get_admin1_regions("peru")[,2]
colnames(wiki_table) <- c("value", "region")

dfn <- data.frame(proof_regions[1:26], wiki_table$region[1:26])

chn_wkd[grep(paste0('[[:space:]]',dfn[17,2]),dfn[,1], ignore.case = T),1]
chn_wkd[11,]

dfn[17,2]
wiki_table[17,]
# grep(paste0('[[:space:]]',dfn[17,2]),dfn[,1], ignore.case = T) %>% length()
# grep(paste0('[[:space:]]',dft[10,2]),dft[,1], ignore.case = T)

#Choropeth-names with wiki data: chn_wkd
chn_wkd <- data.frame("region" = proof_regions, "value" = rep(0,26))

for (i in 1:24){
  if (length(grep(paste0('[[:space:]]',dfn[i,2]),dfn[,1], ignore.case = T)) != 0) {
    ind <- grep(paste0('[[:space:]]',dfn[i,2]),dfn[,1], ignore.case = T)
    chn_wkd[ind,2] <- wiki_table[i,1] 
  }
}
View(chn_wkd)

country.map = admin1.map[admin1.map$admin %in% "peru",]
ggplot(country.map, aes(long, lat, group=group)) + 
    geom_polygon()

admin1_choropleth("peru", data.frame(rep(0,26)))
```

```{r test22}
admin1_map("japan")

data(df_japan_census)
df_japan_census$value = df_japan_census$pop_density_km2_2010
admin1_choropleth("japan", df_japan_census)
```



## 3) Individual Phone Number Maps
Below, you can find an organized list of area code maps done by hand. This list also includes any sources which are in and of themselves complete, in the sense of having an easily-accessible (preferably Wikipedia-based) area code map: 









